comgen-outline

### 2017-12-21 paper/case2-habituals.Rmd
play around with plots, edits to habituals fixed model results

# m.hab.samp <- data.frame()
#  m.hab.fixed.samp <- data.frame()
# 
# for (i in seq(1, n_chains)){
#   # mi <- fread(paste(project.path,  "models/habituals/results/",
#   #                   model_prefix, "smtncs_habitual-",
#   #                   n_samples, "_burn", burn, "_lag", lg, "_chain", i, ".csv", sep = ""))
#   # m.hab.samp <- bind_rows(m.hab.samp, mi %>% mutate(chain = i))
# 
#   # mi.fixed <- fread(paste(project.path,  "models/habituals/results/",
#   #                   model_prefix, "smtncs_some-",
#   #                   n_samples, "_burn", burn, "_lag", lg, "_chain", i, ".csv", sep = ""))
#   mi.fixed <- fread(paste(project.path,  "models/habituals/results/",
#                           "results-habituals-jointModel-inferFixedThresholdCts-S1-smtncs_some-100000_burn50000_lag20_chain", i, ".csv", sep = ""))
#   #
#   m.hab.fixed.samp <- bind_rows(m.hab.fixed.samp, mi.fixed %>% mutate(chain = i))
# }

@@

# save(m.hab.fixed.samp,
#      file = paste(project.path,  "results-habituals-jointModel-inferFixedThresholdCts-noise-S1-smtncs_some-100000_burn50000_lag20_2chains.RData", sep = ""))

@@

# load(paste(project.path,  "models/habituals/results/", model_prefix, "some-",
#                     n_samples, "_burn", burn, "_lag", lg, "_",n_chains , "chains.RData", sep = ""))
# load(paste(project.path,  
#            "results-habituals-jointModel-inferFixedThreshold-S1-smtncs_some-100000_burn50000_lag20_2chains.RData", sep = ""))
load(paste(project.path,  "results-habituals-jointModel-inferFixedThresholdCts-noise-S1-smtncs_some-100000_burn50000_lag20_2chains.RData", sep = ""))

@@

fixed.hab.params.posterior <-  m.hab.fixed.samp %>% 
  filter(B %in% c("noise", "fixedThreshold")) %>%
  group_by(B) %>% 
  summarize(MAP = estimate_mode(val),
            cred_upper = hdi_upper(val),
            cred_lower = hdi_lower(val))
fixed.threshold.hab.posterior <- filter(fixed.hab.params.posterior, B == "fixedThreshold")
fixed.noise.hab.posterior <- filter(fixed.hab.params.posterior, B == "noise")
m.hab.somemodel.endorsement <- m.hab.fixed.samp %>%
  filter(type == 'predictive') %>%
  rename(habitual = B, time_period = D, binned_freq = E) %>%
  group_by(habitual, time_period, binned_freq) %>%
  summarize(MAP = estimate_mode(val),
            cred_upper = hdi_upper(val),
            cred_lower = hdi_lower(val))

@@

habituals.endorsement.bars <- ggplot(md.hab.long, aes ( y = prediction, ymin = prediction_lower, ymax = prediction_upper,
                  x = time_period, fill = logAnnualRate))+
  #geom_abline(intercept = 0, slope = 1, lty = 3)+
  #geom_linerange(alpha = 0.15)+
  geom_col(position = position_dodge(), color = 'black', width = 0.7)+
  geom_errorbar(alpha = 0.7, width = 0.3)+
  #geom_point(shape = 21, size = 3)+
  #scale_x_continuous(limits = c(-0.01, 1.01), breaks = c(0,  1))+
  scale_y_continuous(limits = c(-0.01, 1.01), breaks = c(0, 1))+
  #scale_fill_continuous(low = "#2b83ba", high = "#d7191c")+
  coord_fixed()+
  ylab("Habitual endorsement")+
  xlab("Frequency")+
  facet_grid(src~habitual, scales = 'free')+
  theme(strip.text.y = element_text(hjust = 0, angle = 0),#, vjust = 1),
        axis.text.x = element_text(vjust = 1, hjust = 1, angle = 45),
        legend.position = "bottom") +
   guides(fill = F)
##habituals.endorsement.bar
#fixed.hab.params.posterior
```

@@

We next examine an endorsement model based on a fixed-threshold semantics.
A fixed-threshold model commits the habitual to conveying that a person *does the action with some frequency*, and that frequency is common to all actions.
As in Case Study 1, we incorporate this model into a Bayesian joint-inference model to infer the fixed-threshold and simultaneously predict both the priors data and the endorsement data.
Because we observe no difference between the felicity judgments for habituals of male and female characters, we use a 50\% mixture of the inferred priors for each gender to construct a single frequency distribution $P(h)$.
We assume the referent-prevalence $h'$ being conveyed by the endorsement model (Eq. \ref{eq:S1}) is the frequency provided to participants (e.g., 3 times in the past year).
Additionally, to account for statements that would be literally false under this model (frequencies that fall below the fixed threshold), we include an additional noise parameter, as we did in Case Study 1. 
To learn about the credible values of the model's parameters and generate predictions given those inferred parameter values, we collected 3 MCMC chains of 100,000 iterations, discarding the first 50,000 iterations for burn in.

```{r hab-fixed-parameters}

@@

The fixed-threshold model is ill-defined for this data set.
The data-analytic model infers that `r 100*round(fixed.noise.hab.posterior[[1,"MAP"]], 2)`% [`r 100*round(fixed.noise.hab.posterior[[1,"cred_lower"]], 2)`,
`r 100*round(fixed.noise.hab.posterior[[1,"cred_upper"]], 2)`] of the data is noise.
The inferred (fixed) threshold in units of number of times per year is `r round(exp(fixed.threshold.hab.posterior[[1,"MAP"]]), 2)` [`r round(exp(fixed.threshold.hab.posterior[[1,"cred_lower"]]), 2)`,
`r round(exp(fixed.threshold.hab.posterior[[1,"cred_upper"]]), 2)`]. 
That is, the best way this model has to deal with the endorsement data is to say there is some arbitrary threshold and a large percentage of deviations are due to noise. 
Thus, our identical data analytic strategy paired with a fixed-threshold model is unable to predict the range of endorsements ($r^2(`r r2.hab.n`) = `r r2.habituals.rsa.fixed`$; MSE=$`r mse.habituals.rsa.fixed`$).^[
  An alternative formulation of this model could fix the threshold to be the lowest possible threshold (analagous to the quantifier "some"). 
  This model would explain even less data than the current model (if it was better, this situation would have been inferred).
  That model would have trouble differentiating between different frequencies for the same habitual statement (as the current model does), analogous to how "some" is equally true of 20\% or 50\%.
]

<!-- The lowest frequency used in our data set was $0.2$ times per year.  -->
<!-- Thus, all of the utterances evaluated under a fixed-threshold model were literally true.  -->

>>>>

# m.hab.samp <- data.frame()
#  m.hab.fixed.samp <- data.frame()

#for (i in seq(1, n_chains)){
  # mi <- fread(paste(project.path,  "models/habituals/results/",
  #                   model_prefix, "smtncs_habitual-",
  #                   n_samples, "_burn", burn, "_lag", lg, "_chain", i, ".csv", sep = ""))
  # m.hab.samp <- bind_rows(m.hab.samp, mi %>% mutate(chain = i))

  # mi.fixed <- fread(paste(project.path,  "models/habituals/results/",
  #                   model_prefix, "smtncs_some-",
  #                   n_samples, "_burn", burn, "_lag", lg, "_chain", i, ".csv", sep = ""))
#   mi.fixed <- fread(paste(project.path,  "models/habituals/results/",
#                           "results-habituals-jointModel-inferFixedThreshold-S1-smtncs_some-100000_burn50000_lag20_chain", i, ".csv", sep = ""))
#   #
#   m.hab.fixed.samp <- bind_rows(m.hab.fixed.samp, mi.fixed %>% mutate(chain = i))
# }

@@

# save(m.hab.fixed.samp,
#      file = paste(project.path,  "models/habituals/results/results-habituals-jointModel-inferFixedThreshold-noise-S1-smtncs_some-100000_burn50000_lag20_2chains.RData", sep = ""))

@@

# load(paste(project.path,  "models/habituals/results/", model_prefix, "some-",
#                     n_samples, "_burn", burn, "_lag", lg, "_",n_chains , "chains.RData", sep = ""))
load(paste(project.path,
           "models/habituals/results/results-habituals-jointModel-inferFixedThreshold-noise-S1-smtncs_some-100000_burn50000_lag20_2chains.RData", sep = ""))
# load(paste(project.path,  "results-habituals-jointModel-inferFixedThresholdCts-noise-S1-smtncs_some-100000_burn50000_lag20_2chains.RData", sep = ""))

@@

fixed.hab.params.posterior <-  m.hab.fixed.samp %>% 
  filter(B %in% c("noise", "fixedThreshold")) %>%
  group_by(B) %>% 
  summarize(MAP = estimate_mode(val),
            cred_upper = hdi_upper(val),
            cred_lower = hdi_lower(val))

fixed.threshold.hab.posterior <- filter(fixed.hab.params.posterior, B == "fixedThreshold")
fixed.noise.hab.posterior <- filter(fixed.hab.params.posterior, B == "noise")

m.hab.somemodel.endorsement <- m.hab.fixed.samp %>%
  filter(type == 'predictive') %>%
  rename(habitual = B, time_period = D, binned_freq = E) %>%
  group_by(habitual, time_period, binned_freq) %>%
  summarize(MAP = estimate_mode(val),
            cred_upper = hdi_upper(val),
            cred_lower = hdi_lower(val))

@@

habituals.endorsement.bars <- ggplot(md.hab.long, aes ( y = prediction, ymin = prediction_lower, ymax = prediction_upper,
                  x = time_period, fill = logAnnualRate))+
  #geom_abline(intercept = 0, slope = 1, lty = 3)+
  #geom_linerange(alpha = 0.15)+
  geom_col(position = position_dodge(), color = 'black', width = 0.7)+
  geom_errorbar(alpha = 0.7, width = 0.3)+
  #geom_point(shape = 21, size = 3)+
  #scale_x_continuous(limits = c(-0.01, 1.01), breaks = c(0,  1))+
  scale_y_continuous(limits = c(-0.01, 1.01), breaks = c(0, 1))+
  #scale_fill_continuous(low = "#2b83ba", high = "#d7191c")+
  coord_fixed()+
  ylab("Habitual endorsement")+
  xlab("Frequency")+
  facet_grid(src~habitual, scales = 'free')+
  theme(strip.text.y = element_text(hjust = 0, angle = 0),#, vjust = 1),
        axis.text.x = element_text(vjust = 1, hjust = 1, angle = 45),
        legend.position = "bottom") +
   guides(fill = F)
#habituals.endorsement.bars
#fixed.hab.params.posterior
#fixed.hab.params.posterior
```

@@

We next examine an endorsement model based on a fixed-threshold semantics.
A fixed-threshold model commits the habitual to conveying that a person *does the action with some frequency*, and that frequency is common to all actions.
As in Case Study 1, we incorporate this model into a Bayesian joint-inference model to infer the fixed-threshold and simultaneously predict both the priors data and the endorsement data.
Because we observe no difference between the felicity judgments for habituals of male and female characters, we use a 50\% mixture of the inferred priors for each gender to construct a single frequency distribution $P(p_f)$, which we use for both fixed-threshold and uncertain-threshold models.
We assume the referent-prevalence $h'$ being conveyed by the endorsement model (Eq. \ref{eq:S1}) is the frequency provided to participants (e.g., 3 times in the past year).
Additionally, to account for statements that would be literally false under this model (frequencies that fall below the fixed threshold), we include an additional noise parameter, as we did in Case Study 1. 
To learn about the credible values of the model's parameters and generate predictions given those inferred parameter values, we collected 3 MCMC chains of 100,000 iterations, discarding the first 50,000 iterations for burn in.

```{r hab-fixed-parameters}
```

@@

The fixed-threshold model exhibits a small dynamic range of endorsements. 
The data analytic model infers that a low threshold is likely: the inferred threshold in units of number of times per year is `r round(exp(fixed.threshold.hab.posterior[[1,"MAP"]]), 2)` [`r round(exp(fixed.threshold.hab.posterior[[1,"cred_lower"]]), 2)`,
`r round(exp(fixed.threshold.hab.posterior[[1,"cred_upper"]]), 2)`].
The lowest frequency used in our data set was $0.6$ times per year (3 times every 5 years).
Thus, all of the utterances evaluated under this fixed-threshold model were literally true.
Therefore, the lowest endorsement this model can apply to an utterance is 0.5 (since both the habitual and silence are always true).

Differences in endorsement rate are due to the informativeness of the fixed-threshold model.
This varies by items, but not by frequency within an item, analogous to how "some" is equally true of 20\% or 50\%.

The data-analytic model infers that `r 100*round(fixed.noise.hab.posterior[[1,"MAP"]], 2)`% [`r 100*round(fixed.noise.hab.posterior[[1,"cred_lower"]], 2)`,
`r 100*round(fixed.noise.hab.posterior[[1,"cred_upper"]], 2)`] of the data is noise.

<!-- That is, the best way this model has to deal with the endorsement data is to say there is some arbitrary threshold and a large percentage of deviations are due to noise.  -->
<!-- Thus, our identical data analytic strategy paired with a fixed-threshold model is unable to predict the range of endorsements ($r^2(`r r2.hab.n`) = `r r2.habituals.rsa.fixed`$; MSE=$`r mse.habituals.rsa.fixed`$).^[ -->
<!--   An alternative formulation of this model could fix the threshold to be the lowest possible threshold (analagous to the quantifier "some").  -->
<!--   This model would explain even less data than the current model (if it was better, this situation would have been inferred). -->
<!--   That model would have trouble differentiating between different frequencies for the same habitual statement (as the current model does), analogous to how "some" is equally true of 20\% or 50\%. -->
<!-- ] -->

<!-- The fixed-threshold model is ill-defined for this data set. -->
<!-- The data-analytic model infers that `r 100*round(fixed.noise.hab.posterior[[1,"MAP"]], 2)`% [`r 100*round(fixed.noise.hab.posterior[[1,"cred_lower"]], 2)`, -->
<!-- `r 100*round(fixed.noise.hab.posterior[[1,"cred_upper"]], 2)`] of the data is noise. -->
<!-- The inferred (fixed) threshold in units of number of times per year is `r round(exp(fixed.threshold.hab.posterior[[1,"MAP"]]), 2)` [`r round(exp(fixed.threshold.hab.posterior[[1,"cred_lower"]]), 2)`, -->
<!-- `r round(exp(fixed.threshold.hab.posterior[[1,"cred_upper"]]), 2)`].  -->
<!-- That is, the best way this model has to deal with the endorsement data is to say there is some arbitrary threshold and a large percentage of deviations are due to noise.  -->
<!-- Thus, our identical data analytic strategy paired with a fixed-threshold model is unable to predict the range of endorsements ($r^2(`r r2.hab.n`) = `r r2.habituals.rsa.fixed`$; MSE=$`r mse.habituals.rsa.fixed`$).^[ -->
<!--   An alternative formulation of this model could fix the threshold to be the lowest possible threshold (analagous to the quantifier "some").  -->
<!--   This model would explain even less data than the current model (if it was better, this situation would have been inferred). -->
<!--   That model would have trouble differentiating between different frequencies for the same habitual statement (as the current model does), analogous to how "some" is equally true of 20\% or 50\%. -->
<!-- ] -->
